---
title: "EDRM 718 Adam Brown Final Project"
author: "Adam Brown"
date: "`r Sys.Date()`"
format: html
editor: visual
---

```{r}
#| include: false

library(here)
library(gt)
library(readr)
library(readxl)
library(tibble)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)

```

```{r}
#| include: false


# First, I'm reading in the Excel files using read_excel so I can then convert them to CSV files. I could save each file in Excel as a CSV but why not let R do this for me. Then I will save the CSV files in my data folder so that I can import them using read_csv.

# Reading in the Excel files into R.

job_prep_demographics <- read_excel(here("data", "Job Prep Demographics.xlsx"))
job_prep_gpa_data <- read_excel(here("data", "Job Prep GPA Data.xlsx"))
job_prep_survey_results <- read_excel(here("data", "Job Prep Survey Results.xlsx"))

# Writing the data to CSV files in the 'data' folder.

write_csv(job_prep_demographics, here("data", "Job_Prep_Demographics.csv"))
write_csv(job_prep_gpa_data, here("data", "Job_Prep_GPA_Data.csv"))
write_csv(job_prep_survey_results, here("data", "Job_Prep_Survey_Results.csv"))


```

```{r}
#| include: false


# Reading the CSV files into R using read_csv..

job_prep_demographics <- read_csv(here("data", "Job_Prep_Demographics.csv"), col_names = FALSE)
job_prep_gpa_data <- read_csv(here("data", "Job_Prep_GPA_Data.csv"), col_names = FALSE)
job_prep_survey_results <- read_csv(here("data", "Job_Prep_Survey_Results.csv"), col_names = FALSE)

# View the first few rows of each dataset to confirm they loaded correctly.

head(job_prep_demographics)
head(job_prep_gpa_data)
head(job_prep_survey_results)


```

```{r}
#| include: false
# Checking the Demo data for any problems that occurred during import.

demo_issues <- problems(job_prep_demographics)

# Viewing the problems.

if (nrow(demo_issues) > 0) {
  print("Problems were found during import:")
  print(import_issues)
} else {
  print("No problems were found during import.")
}

# Checking the gpa data for any problems that occurred during import.

gpa_issues <- problems(job_prep_gpa_data)

# Viewing the problems.

if (nrow(gpa_issues) > 0) {
  print("Problems were found during import:")
  print(import_issues)
} else {
  print("No problems were found during import.")
}

# Checking the survey data for any problems that occurred during import.

survey_issues <- problems(job_prep_survey_results)

# Viewing the problems, if any.

if (nrow(survey_issues) > 0) {
  print("Problems were found during import:")
  print(import_issues)
} else {
  print("No problems were found during import.")
}

```

## Description

This research report explores the relationships between optimisim scores (from the survey data and measures their outlook on employment and career readiness), reported GPA's, and the college students academic disciplines. More specifically, this report seeks to answer these three questions, Is senior-year grade point average (GPA) related to the degree of optimism about future employment? Is the relationship of GPA and optimism consistent or different for different categories of disciplines? Is the level of optimism different for different categories of disciplines? The three datasets include survey results from over 700 college seniors. graduating this spring, representing 23 major universities in the United States. These datasets include demographic data, academic performance data, and data from a short (14-item) Likert-type survey which sought to determine their optimism regarding their readiness and likelihood of obtaining a job after graduation

**Source of Data:** Real World Solutions, Job Prep Demographics, Job Prep GPA, and Job Prep Survey Results. Sourced 24 April 2025.

## Analysis

I begin my analysis reading in the three separate data files, tidying the data that is missing or has problems, and then joining the files together using my unique key, which in this case is Student ID number. Now that I have completed the data processing and cleaning, I start by examining the reported GPA's for the students from the first and second semester. We need to verify consistency of GPA's before we can create a weighted GPA average.

```{r}
#| include: false

# File # 1 - Demo Data.

# I am now doing several things. First, I have to add a column and put the subject in rows and not simply at the top of each section. Second, I am removing the first six rows as I do not need this data and setting the new row one as the header. Third, I am correcting the header value of Humanities that I created in my first two steps by way of processing and I am setting this column header as Subjects. Next, I am setting the column types and removing any Subjects rows that have Other as this was my term for unnecessary data. Lastly, I'm cleaning the column names to valid and consistent.

# Add the discipline column to mark each section (e.g., Humanities, Social Sciences)
job_prep_demographics <- job_prep_demographics %>%
  mutate(discipline = case_when(
    row_number() <= 145 ~ "Humanities",  # Adjust row numbers for Humanities
    row_number() > 150 & row_number() <= 303 ~ "Social Sciences",  # Adjust for Social Sciences
    row_number() > 308 & row_number() <= 463 ~ "Natural Sciences",  # Adjust for Natural Sciences
    row_number() > 468 & row_number() <= 617 ~ "Formal Sciences",  # Adjust for Formal Sciences
    row_number() > 622 ~ "Applied Sciences",  # Adjust for Applied Sciences
    TRUE ~ "Other"
  ))

# Remove the first six rows (metadata/incorrect headers)
job_prep_demographics_clean <- job_prep_demographics %>%
  slice(-1, -2, -3, -4, -5, -6)  # Remove the first 6 rows with non-data

# Set correct column names, where row 1 should be the header
colnames(job_prep_demographics_clean) <- job_prep_demographics_clean[1, ]
job_prep_demographics_clean <- job_prep_demographics_clean[-1, ]  # Remove the first row now that it is the header

# Rename the "Humanities" column to "Subjects"
job_prep_demographics_clean <- job_prep_demographics_clean %>%
  rename(Subjects = Humanities)

# Set the correct column types
job_prep_demographics_clean <- job_prep_demographics_clean %>%
  mutate(
    ID = as.integer(ID),
    `School Code` = as.integer(`School Code`),
    Age = as.integer(Age),
    Gender = as.character(Gender),  # Convert gender to character for easier manipulation
    Subjects = as.character(Subjects)  # Ensure Subjects is a character column
  )

# Remove rows where the value in the Subjects column is "Other"
job_prep_demographics_clean <- job_prep_demographics_clean %>%
  filter(Subjects != "Other")

# Clean column names to make them syntactically valid and consistent
colnames(job_prep_demographics_clean) <- make.names(colnames(job_prep_demographics_clean))

# View the cleaned data
head(job_prep_demographics_clean)
```

```{r}
#| include: false

# File 2 - GPA data.

# Remove the first four rows with non-data and metadata
job_prep_gpa_data_tidy <- job_prep_gpa_data %>%
  slice(-1, -2, -3, -4)  # Remove first four rows

# Set the column names using row 1 (now row 5 after removing the first 4 rows)
colnames(job_prep_gpa_data_tidy) <- job_prep_gpa_data_tidy[1, ]
job_prep_gpa_data_tidy <- job_prep_gpa_data_tidy[-1, ]  # Remove the first row which is now the header

# Rename columns and calculate average GPA
job_prep_gpa_data_tidy <- job_prep_gpa_data_tidy %>%
  rename(
    first_semester_gpa = "S1 GPA",
    second_semester_gpa = "S2 GPA",
    credits_first_semester = "S1 Credits",
    credits_second_semester = "S2 Credits"
  ) 

# Set the column types and format GPA columns
job_prep_gpa_data_tidy <- job_prep_gpa_data_tidy %>%
  # Convert columns to the correct types
  mutate(
    ID = as.integer(ID),
    `School Code` = as.integer(`School Code`),
    Age = as.integer(Age),
    Gender = as.character(Gender),
    credits_first_semester = as.numeric(credits_first_semester),
    first_semester_gpa = as.numeric(first_semester_gpa),
    credits_second_semester = as.numeric(credits_second_semester),
    second_semester_gpa = as.numeric(second_semester_gpa)
  ) %>%
  
  # Round GPA columns to 2 decimal places
mutate(
  first_semester_gpa = round(first_semester_gpa, 2),  # Round to 2 decimal places
  second_semester_gpa = round(second_semester_gpa, 2)  # Round to 2 decimal places
)

# View the tidied data
head(job_prep_gpa_data_tidy)


```

```{r}
#| warning: false
#| include: false

# File 3 - Survey data.

# Slice off the first three rows
job_prep_results_clean <- job_prep_survey_results %>%
  slice(-1, -2, -3)  # Remove first 3 rows

# Set the correct column names (the 4th row, now the first row)
colnames(job_prep_results_clean) <- job_prep_results_clean[1, ]
job_prep_results_clean <- job_prep_results_clean[-1, ]  # Remove the first row now that it is the header

# Rename columns to replace spaces with underscores (for Item 1, Item 2, ...)
colnames(job_prep_results_clean)[2:15] <- gsub(" ", "_", colnames(job_prep_results_clean)[2:15])

# Convert all columns to integer type
job_prep_results_clean <- job_prep_results_clean %>%
  mutate(across(everything(), as.integer))

# View the cleaned dataset
head(job_prep_results_clean)

```

```{r}
#| warning: false
#| include: false

# Check the column names in each dataset to ensure 'ID' is consistent.

colnames(job_prep_demographics_clean)
colnames(job_prep_gpa_data_tidy)
colnames(job_prep_results_clean)

# Check the first few rows of each dataset to make sure 'ID' is present and formatted correctly.

head(job_prep_demographics_clean)
head(job_prep_gpa_data_tidy)
head(job_prep_results_clean)

# Remove any leading or trailing spaces in column names.

colnames(job_prep_demographics_clean) <- trimws(colnames(job_prep_demographics_clean))
colnames(job_prep_gpa_data_tidy) <- trimws(colnames(job_prep_gpa_data_tidy))
colnames(job_prep_results_clean) <- trimws(colnames(job_prep_results_clean))

# Step 4: Ensure 'ID' is an integer in all datasets.

job_prep_demographics_clean <- job_prep_demographics_clean %>%
  mutate(ID = as.integer(ID))

job_prep_gpa_data_tidy <- job_prep_gpa_data_tidy %>%
  mutate(ID = as.integer(ID))

job_prep_results_clean <- job_prep_results_clean %>%
  mutate(ID = as.integer(ID)) 

# Perform the left joins again after ensuring column consistency.

job_prep_combined <- job_prep_demographics_clean %>%
  left_join(
    job_prep_gpa_data_tidy %>%
      select(ID, credits_first_semester, credits_second_semester, first_semester_gpa, second_semester_gpa),
    by = "ID"
  ) %>%
  left_join(
    job_prep_results_clean %>%
      select(ID, Item_1, Item_2, Item_3, Item_4, Item_5, Item_6, Item_7, Item_8, Item_9, Item_10, Item_11, Item_12, Item_13, Item_14),  # Select specific columns from Table 3
    by = "ID"
  )

# View the combined data.

head(job_prep_combined)

```

```{r}
#| warning: false
#| include: false

# Here I am taking my clean joined table and removing the rows that contain NA for survey responses. 

# Remove rows with any NA values.

job_prep_combined_clean <- job_prep_combined %>%
  drop_na()

# Convert columns to the correct types after dropping NA values.

job_prep_combined_clean <- job_prep_combined_clean %>%
  mutate(
    
    # Convert 'Item_1' through 'Item_14' to integers.
    
    across(starts_with("Item_"), as.integer),
    
    
    # Convert GPA and credits columns to numeric.
    
    second_semester_gpa = as.numeric(second_semester_gpa),
    first_semester_gpa = as.numeric(first_semester_gpa),
    credits_first_semester = as.numeric(credits_first_semester),
    credits_second_semester = as.numeric(credits_second_semester)
  )

# View the cleaned data with the updated column types.

head(job_prep_combined_clean)

# View the cleaned data with Item_1 through Item_14 as integers.

head(job_prep_combined_clean)

```

In Figure 1, I created a scatterplot to verify consistency by observing if there a strong linear relationship exists between first and second semester GPA's. In Figure 1, we observe that as the GPA for the first semester increases, so does the GPA for the second semester. This tell us that a strong linear relationship does exist. Since a strong linear relationship exists this GPA's are comparable which means we can create our next scatterplot which will helps us to know if we can predict second semester GPA from first semester GPA.

#### Figure 1. Scatterplot of First Semester GPA Vs Second Semester GPA.

```{r}
#| echo: false
#| warning: false

# Plot the first semester GPA against the second semester GPA.

ggplot(job_prep_combined_clean, aes(x = first_semester_gpa, y = second_semester_gpa)) +
  geom_point() +  
  geom_smooth(method = "lm", col = "red") +  # Add a linear regression line
  labs(title = "First Semester GPA vs Second Semester GPA",
       x = "First Semester GPA",
       y = "Second Semester GPA")


```

In Figure 2, I created a scatterplot to identify if patterns exist or not. We do not observe any shapes or patters in the data, which is good because this means the residuals are randomly distributed. This scatterplot helps me to know that my linear model is a good fit and there shouldn't be significant bias in predicting second semester GPA from first semester. This gives me confidence to create a weighted GPA but first I need to identify and exclude those students who are among those with the ten percent largest residuals. I run this analysis next in Figure 3.

#### Figure 2. Scatterplot of Residuals from Predicted Second Semester GPA.

```{r}
#| echo: false
#| warning: false

# Linear model to predict second semester GPA from first semester GPA.

model <- lm(second_semester_gpa ~ first_semester_gpa, data = job_prep_combined_clean)

# Calculate residuals.

job_prep_combined_clean$residuals <- residuals(model)

# Plot the residuals.

ggplot(job_prep_combined_clean, aes(x = first_semester_gpa, y = residuals)) +
  geom_point() +
  labs(title = "Residuals from Predicted Second Semester GPA",
       x = "First Semester GPA",
       y = "Residuals")


```

```{r}
#| warning: false
#| include: false

# Calculate the 90th percentile of the absolute residuals.

residual_threshold <- quantile(abs(job_prep_combined_clean$residuals), 0.9)

# Exclude students whose absolute residuals are above the 90th percentile.

job_prep_combined_clean_filtered <- job_prep_combined_clean %>%
  filter(abs(residuals) <= residual_threshold)

# Check the number of students excluded.

nrow(job_prep_combined_clean) - nrow(job_prep_combined_clean_filtered)

```

In Figure 3, we can observe the students who are among those with the ten percent largest residuals as they are noted in red. We note from our data that there are 54 students who fall into this category. I will remove those students from the data prior to calculating weighted GPA and we will continue to move through the analysis.

#### Figure 3. Scatterplot of First Semester GPA Vs Second Semester GPA with Large Residuals.

```{r}
#| echo: false
#| warning: false

# Highlight students with the largest residuals in the scatter plot.

ggplot(job_prep_combined_clean, aes(x = first_semester_gpa, y = second_semester_gpa)) +
  geom_point(aes(color = abs(residuals) > residual_threshold)) +  
  geom_smooth(method = "lm", col = "red") +
  labs(title = "First Semester GPA vs Second Semester GPA with Large Residuals Highlighted",
       x = "First Semester GPA",
       y = "Second Semester GPA") +
  scale_color_manual(values = c("black", "red"), labels = c("Normal", "Large Residuals"))

```

```{r}
#| warning: false
#| include: false

# Ensure columns are numeric before calculating the weighted GPA.

job_prep_combined_clean_filtered <- job_prep_combined_clean_filtered %>%
  mutate(
    first_semester_gpa = as.numeric(first_semester_gpa),
    credits_first_semester = as.numeric(credits_first_semester),
    second_semester_gpa = as.numeric(second_semester_gpa),
    credits_second_semester = as.numeric(credits_second_semester)
  )

# Calculate the weighted average GPA.

job_prep_combined_clean_filtered <- job_prep_combined_clean_filtered %>%
  mutate(
    weighted_avg_gpa = (first_semester_gpa * credits_first_semester + second_semester_gpa * credits_second_semester) /
                       (credits_first_semester + credits_second_semester)
  )

# View the updated data with the weighted average GPA.

head(job_prep_combined_clean_filtered)

```

```{r}
#| warning: false
#| include: false

# Call the function from the functions folder.

source("functions/likert_functions.R")

# Using the function to reverse score the below items.

job_prep_combined_clean_filtered <- calculate_likert_score(
  job_prep_combined_clean_filtered,   # The data
  scale = 6,                          # The scale used for the items (e.g., 6-point scale)
  reverse_items = c(4, 9, 12, 13)         # Items to reverse score
)

# View the data with the total score.

head(job_prep_combined_clean_filtered)

```

Question #1

```{r}
#| warning: false
#| include: false

# Call the function from the functions folder.

source("functions/find_largest_residuals.R")

# Example usage of the function with 'weighted_avg_gpa' as the explanatory variable.

result <- find_largest_residuals(
  job_prep_combined_clean_filtered,     # The data frame
  response_var = "total_score",          # The response variable
  explanatory_var = "weighted_avg_gpa", # The explanatory variable (e.g., GPA)
  p = 10                                 # Top 10 largest residuals
)

# Add the result as a new column in the data frame and rename the table.

job_prep_combined_clean_filtered_largest <- job_prep_combined_clean_filtered %>%
  mutate(large_residuals = result)

# View the data with the new column indicating large residuals.

head(job_prep_combined_clean_filtered_largest)

```

In figure 4, we start to answer our first research question, Is senior-year grade point average (GPA) related to the degree of optimism about future employment? We start by verifying consistency by reviewing the ten students with the largest residuals in the dataset. We can observe these students by the red dots in Figure 4. After reviewing these students in the dataset, they show no pattern between them on Subjects, GPA, Optimism Score, etc. So, it will not stregthen our model to remove these students, thus we retain them in the dataset.

#### Figure 4. Scatterplot of Residuals of Senior Year GPA Vs Optimism Score.

```{r}
#| echo: false
#| warning: false


# Add residuals to the data frame using weighted_avg_gpa.

job_prep_combined_clean_filtered_largest <- job_prep_combined_clean_filtered_largest %>%
  mutate(residuals = residuals(lm(total_score ~ weighted_avg_gpa, data = job_prep_combined_clean_filtered_largest)))

# Create the plot for residuals vs weighted_avg_gpa.

ggplot(job_prep_combined_clean_filtered_largest, aes(x = weighted_avg_gpa, y = residuals)) +
  geom_point(aes(color = large_residuals), size = 2) +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  # Line at residual = 0
  labs(title = "Residuals vs Senior Year GPA", 
       x = "Senior Year GPA", y = "Residuals") +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "blue")) +  # Color large residuals in red
  theme_minimal()

```

#### Figure 5. Histogram of the Frequency of Absolute Residuals.

```{r}
#| echo: false
#| warning: false


# Create a histogram of absolute residuals.
ggplot(job_prep_combined_clean_filtered_largest, aes(x = abs(residuals))) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  geom_vline(aes(xintercept = sort(abs(residuals), decreasing = TRUE)[10]), 
             color = "red", linetype = "dashed") +
  labs(title = "Histogram of Absolute Residuals", 
       x = "Absolute Residuals", y = "Frequency") +
  theme_minimal()

```

#### Figure 6. Scatterplot of Senior Year GPA (Weighted GPA) Vs Optimism Score.

```{r}
#| echo: false
#| warning: false

ggplot(job_prep_combined_clean_filtered_largest, aes(x = weighted_avg_gpa, y = total_score)) +
  geom_point() +  # Scatter plot of the data points
  geom_smooth(method = "lm", col = "red") +  # Add a linear regression line
  labs(title = "Senior-Year GPA vs Optimism Score", 
       x = "Weighted Average GPA", y = "Optimism Score")



```

#### Table 1. Correlation Between Weighted GPA and Optimism Score.

```{r}
#| echo: false
#| warning: false


# Calculate the correlation between weighted_avg_gpa and total_score.

correlation <- cor(job_prep_combined_clean_filtered_largest$weighted_avg_gpa, 
                   job_prep_combined_clean_filtered_largest$total_score, 
                   use = "complete.obs")

# Round the correlation coefficient to two decimal places.

correlation <- round(correlation, 2)

# Create the tibble.

correlation_tibble <- tibble(
  "Variable 1" = "weighted_avg_gpa",
  "Variable 2" = "total_score",
  Correlation_Coefficient = correlation
)

correlation_tibble %>%
  gt() %>%
  tab_header(title = "Correlation Between Weighted GPA and Optimism Score") %>%
  fmt_number(
    columns = vars(Correlation_Coefficient),
    decimals = 2
  )


```

#### Table 2. Linear Regression of Optimism Score and Weighted GPA.

```{r}
#| echo: false
#| warning: false

# Run a linear regression model.

model <- lm(total_score ~ weighted_avg_gpa, data = job_prep_combined_clean_filtered_largest)

# Summary of the regression model.

model_summary <- summary(model)

# Extract coefficients, p-values, and other relevant results.

regression_results <- tibble(
  Term = rownames(model_summary$coefficients),
  Estimate = round(model_summary$coefficients[, 1], 2),
  Std_Error = round(model_summary$coefficients[, 2], 2),
  t_Value = round(model_summary$coefficients[, 3], 2),
  p_Value = round(model_summary$coefficients[, 4], 2)
)

regression_results %>%
  gt() %>%
  tab_header(
    title = "Linear Regression Results: Optimism Score vs Weighted GPA"
  ) %>%
  fmt_number(
    columns = vars(Estimate, Std_Error, t_Value, p_Value),
    decimals = 2
  ) %>%
  cols_label(
    Term = "Term",
    Estimate = "Estimate",
    Std_Error = "Standard Error",
    t_Value = "t-Statistic",
    p_Value = "p-Value"
  )

```

Question #2

#### Figure 6. Scatterplot of Weighted GPA Vs Optimism Score by Subject.

```{r}
#| echo: false
#| warning: false

# Scatter plot of weighted_avg_gpa vs total_score for each subject category.

ggplot(job_prep_combined_clean_filtered_largest, aes(x = weighted_avg_gpa, y = total_score)) +
  geom_point(aes(color = Subjects)) +  # Color by subject category
  geom_smooth(method = "lm", aes(color = Subjects), se = FALSE) +  
  labs(title = "Relationship Between Weighted Avg GPA and Optimism Score by Subject",
       x = "Weighted Average GPA", y = "Optimism Score") +
  theme_minimal()

```

#### Table 3. Linear Regression of Subjects.

```{r}
#| echo: false
#| warning: false

# Fit separate linear models for each subject category.

model_by_subject <- job_prep_combined_clean_filtered_largest %>%
  group_by(Subjects) %>%
  do(model = lm(total_score ~ weighted_avg_gpa, data = .))

# Extract coefficients and p-values from each model, round to two decimal places.

model_summary <- model_by_subject %>%
  summarise(
    Subjects = first(Subjects),  
    Estimate = round(coef(model)[2], 2),  
    P_value = round(summary(model)$coefficients[2, 4], 2)  
  )


model_summary %>%
  gt() %>%
  tab_header(
    title = "Linear Regression Model Results by Subject"
  ) %>%
  fmt_number(
    columns = vars(Estimate, P_value),
    decimals = 2
  ) %>%
  cols_label(
    Subjects = "Subject",  # The 'Subjects' column should now work correctly
    Estimate = "Coefficient of Weighted GPA",
    P_value = "p-value"
  )


```

#### Table 4. Interaction Model of Optimism Score, Weighted GPA, and Subjects.

```{r}
#| echo: false
#| warning: false

# Fit an interaction model.

interaction_model <- lm(total_score ~ weighted_avg_gpa * Subjects, data = job_prep_combined_clean_filtered_largest)

# Summary of the interaction model.

interaction_summary <- summary(interaction_model)

# Extract coefficients, standard errors, t-values, and p-values from the interaction model.

interaction_results <- tibble(
  Term = rownames(interaction_summary$coefficients),
  Estimate = round(interaction_summary$coefficients[, 1], 2),
  Std_Error = round(interaction_summary$coefficients[, 2], 2),
  t_Value = round(interaction_summary$coefficients[, 3], 2),
  p_Value = round(interaction_summary$coefficients[, 4], 2)
)

interaction_results %>%
  gt() %>%
  tab_header(
    title = "Interaction Model Results: Optimism Score vs Weighted GPA and Subjects"
  ) %>%
  fmt_number(
    columns = vars(Estimate, Std_Error, t_Value, p_Value),
    decimals = 2
  ) %>%
  cols_label(
    Term = "Term",
    Estimate = "Estimate",
    Std_Error = "Standard Error",
    t_Value = "t-Statistic",
    p_Value = "p-Value"
  )


```

Question #3

#### Figure 7. Distribution of Optimism Scores Across Subjects.

```{r}
#| echo: false
#| warning: false

# Box plot to visualize the distribution of optimism_score across subjects.

ggplot(job_prep_combined_clean_filtered_largest, aes(x = Subjects, y = total_score)) +
  geom_boxplot(aes(fill = Subjects)) +  # Box plot colored by subject
  labs(title = "Distribution of Optimism Score Across Subjects", 
       x = "Subjects", y = "Optimism Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#### Table 5. One-way ANOVA of Optimism Score Across Subjects.

```{r}
#| echo: false
#| warning: false

# Conduct a one-way ANOVA to compare total_score across subjects.

anova_results <- aov(total_score ~ Subjects, data = job_prep_combined_clean_filtered_largest)

# View the summary of the ANOVA results.

anova_summary <- summary(anova_results)

# Extract the relevant results from the ANOVA summary.

anova_tibble <- tibble(
  Df = anova_summary[[1]]$`Df`,  # Degrees of freedom
  Sum_Sq = round(anova_summary[[1]]$`Sum Sq`, 2),  
  Mean_Sq = round(anova_summary[[1]]$`Mean Sq`, 2),  
  F_value = round(anova_summary[[1]]$`F value`, 2),  
  Pr_F = round(anova_summary[[1]]$`Pr(>F)`, 2)  
)

# Display the ANOVA results using gt for a table.

anova_results_gt <- anova_tibble %>%
  gt() %>%
  tab_header(
    title = "ANOVA Results for Optimism Score by Subject"
  ) %>%
  cols_label(
    Df = "Degrees of Freedom",
    Sum_Sq = "Sum of Squares",
    Mean_Sq = "Mean Squares",
    F_value = "F Value",
    Pr_F = "p-value"
  ) %>%
  fmt_number(
    columns = vars(Sum_Sq, Mean_Sq, F_value, Pr_F),
    decimals = 2
  )

# Print the table.

anova_results_gt

```

#### Table 6. ?.

```{r}
#| echo: false
#| warning: false

# Perform Tukey's HSD post-hoc test.

tukey_results <- TukeyHSD(anova_results)

# Extract the Tukey results for 'Subjects' comparison.

tukey_df <- as.data.frame(tukey_results$`Subjects`)

# Convert to tibble and round the values to 2 decimal places.

tukey_tibble <- tibble(
  Comparison = rownames(tukey_df),  
  Difference = round(tukey_df$diff, 2),
  `Lower Bound` = round(tukey_df$lwr, 2),
  `Upper Bound` = round(tukey_df$upr, 2),
  p_value = round(tukey_df$`p adj`, 2)  
)

tukey_results_gt <- tukey_tibble %>%
  gt() %>%
  tab_header(
    title = "Tukey's HSD Post-Hoc Test Results"
  ) %>%
  cols_label(
    Comparison = "Comparison of Subjects",
    Difference = "Difference in Means",
    `Lower Bound` = "Lower Bound of CI",
    `Upper Bound` = "Upper Bound of CI",
    p_value = "p-value"
  ) %>%
  fmt_number(
    columns = vars(Difference, `Lower Bound`, `Upper Bound`, p_value),
    decimals = 2
  )

# Print the Tukey results table.

tukey_results_gt
```

## Conclusion
